Hydra Battles and Cie

### 2020-12-03 doc/chapter-powers.tex
minor revisions on addition chains part (.tex, .v)

\subsubsection{Removing duplicate computations}

@@

By induction\footnote{Simple induction, since $n$ has type \texttt{nat}.}
 over $n$, we obtain the following equality.

 \begin{Coqsrc}
Lemma fib_power n : fib n = let (a,b) := power (M:=Mul2) (1,0) n
                            in (a+b).   
\end{Coqsrc}

Thus, any function able to compute more or less efficiently powers in a monoid will
give an algorithm for computing Fibonacci numbers. Unlike the \emph{ad-hoc} aforementionned proofs of~\cite{BC04}, the correctness of such an algorithm is a direct consequence 
of the correctness of the used powering function.
Several examples will be presented in the rest of this document.

@@

@@

Definition chain_correct (p:positive) (c: chain) :=
  chain_correct_nat c (Pos.to_nat p). 
\end{Coqsrc}

@@

\begin{Coqanswer}
...
============================
   a * a * a == a * (a * (a * E_one))}
\end{Coqanswer}

@@

@@

@@

>>>>

\subsubsection{Removing duplicate computations}
\label{sect:fibonacci-mul2}

@@

By induction\footnote{Simple induction, since $n$ has type \texttt{nat}.}
 over $n$, we obtain a the following equalities.

\begin{Coqsrc}
Definition fib_mul2 n := let (a,b) := power (M:=Mul2) (1,0) n   in (a+b).

Lemma fib_mul2_OK_0 (n:nat) : power (M:=Mul2) (1,0) (S (S n)) =
                              (fib (S n), fib n).

Lemma fib_mul2_OK n : fib n = fib_mul2 n.
\end{Coqsrc}

Thus, any function able to compute more or less efficiently powers in a monoid will
give an algorithm for computing Fibonacci numbers. Unlike the \emph{ad-hoc} aforementionned proofs of~\cite{BC04}, the correctness of such an algorithm is a direct consequence 
of the correctness of the used powering function.
Several examples will be presented in the rest of this document
(in sections~\vref{sect:fibonacci-pos-bpow}).

@@

\subsection{Fibonacci, once again}
\label{sect:fibonacci-pos-bpow}

We can use the function \texttt{Pos\_bpow} for computing Fibonacci numbers
(see Section~\vref{sect:fibonacci-mul2}).

\begin{Coqsrc}
Definition fib_pos n :=
  let (a,b) := Pos_bpow (M:= Mul2) (1,0) n in
  (a+b).

Time Compute fib_pos 153%positive.
\end{Coqsrc}

\begin{Coqanswer}
  68330027629092351019822533679447
     : N 
Finished transaction in 0.002 secs (0.002u,0.s) (successful)
\end{Coqanswer}

Fibonacci will come back in Sect.~\vref{sect:fibonacci-euclidean}.

@@

Definition chain_correct (p:positive) (c: chain) :=
  chain_correct_nat c (Pos.
to_nat p). 
\end{Coqsrc}

@@

\begin{Coqanswer}
...
============================
a * a * a == a * (a * (a * E_one))}

@@

\begin{Coqsrc}
Function dicho_aux (p:positive) {struct p} : positive :=
 match p with
   | 1%positive   =>  xH
   | 2%positive |   3%positive  => 2 
   | xO (xO q) | xO (xI q) | xI (xO q) | xI (xI q) =>
                                         xO (dicho_aux q)
 end.

Definition dicho  (p:positive) : positive :=
  N2pos (N.div (Npos p) (Npos (dicho_aux p))).

Instance Dicho_strat : Strategy dicho.
\end{Coqsrc}

\subsection{Other strategies}
For comparison's sake, we define two other strategies, much simpler but statically less efficient than the dichotomic strategy.

\emph{From Module~\href{../theories/html/hydras.additions.BinaryStrat.html}{additions.BinaryStrat}.}

\begin{Coqsrc}
Definition half (p:positive) :=
  match p with xH => xH
          |    xI q | xO q =>  q
  end.

Definition two (p:positive) := 2%positive.

Instance Binary_strat : Strategy half.
Proof.
  split; destruct p; unfold half; try lia.
Qed.

Instance Two_strat : Strategy two.
Proof.
  split;unfold two; lia.
Qed.
\end{Coqsrc}

In Sect.~\ref{sect:test-strat}, we compare the three strategies with respect to the lenght of the built chains.

@@

\label{sect:test-strat}

The following tests show various examples of chains for the same exponent, using different strategies. The dichotomic strategy seems clearly to be the winner (at least on this sample).

\begin{Coqsrc}
Compute chain_length (make_chain two 56789).
\end{Coqsrc}

\begin{Coqanswer}
= 25%nat : nat  
\end{Coqanswer}

\begin{Coqsrc}
Compute chain_length (make_chain half 56789).
\end{Coqsrc}

\begin{Coqanswer}
 = 25%nat : nat
\end{Coqanswer}

\begin{Coqsrc}
Compute chain_length (make_chain dicho 56789).
\end{Coqsrc}

\begin{Coqanswer}
= 21%nat : nat 
\end{Coqanswer}

\begin{Coqsrc}
Compute chain_length (make_chain two 3456789).
\end{Coqsrc}

\begin{Coqanswer}
= 33%nat : nat
\end{Coqanswer}

\begin{Coqsrc}
Compute chain_length (make_chain half 3456789).
\end{Coqsrc}

\begin{Coqanswer}
(= 33%nat : nat
\end{Coqanswer}

\begin{Coqsrc}
Compute chain_length (make_chain dicho 3456789).
\end{Coqsrc}

@@

\subsection{Fibonacci, \emph{le retour}}
\label{sect:fibonacci-euclidean}

It is now possible to use euclidean addition chains for computing Fibonacci numbers
(see Sections~\vref{sect:fibonacci-mul2} and~\vref{sect:fibonacci-pos-bpow}).

The following function is parameterized by a startegy $\gamma$.

\begin{Coqsrc}
Definition fib_eucl gamma `{Hgamma: Strategy gamma} n :=
  let c := make_chain gamma  n
  in let r := chain_apply c (M:=Mul2) (1,0) in
       fst r + snd r.

Compute fib_eucl dicho 153.
\end{Coqsrc}

\begin{Coqanswer}
    = 68330027629092351019822533679447
     : N
Finished transaction in 0.002 secs (0.002u,0.s) (successful)
\end{Coqanswer}

\begin{Coqsrc}
Compute fib_eucl two 153.
\end{Coqsrc}

\begin{Coqanswer}
    = 68330027629092351019822533679447
     : N
Finished transaction in 0.003 secs (0.003u,0.s) (successful)
\end{Coqanswer}

\begin{Coqsrc}
Compute fib_eucl half 153.
\end{Coqsrc}

\begin{Coqanswer}
    = 68330027629092351019822533679447
     : N
Finished transaction in 0.003 secs (0.003u,0.s) (successful)
\end{Coqanswer}
